---
title: "Data processing for assignment 2"
author: "Cas Blaauw, Dominic Comerford and Netanja Harlianto"
date: "09/06/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

### Introduction

We chose the Coffee Quality dataset, which has been scraped from the Coffee Quality Institute database in 2018.
Our dataset contains information on two different coffee plants: Arabica and Robusta. Information about quality measures, coffee bean data, and farm data are included. Quality measures include: aroma, flavor, aftertaste, acidity, body, balance, uniformity, cup cleanliness, sweetness, moisture, and defects. Bean data includes the species, color, and the method of processing. Farmdata in which we are interested includes region/country of origin, and the altitude at which the farm is located with respect to sea level.

### Preparatory work

```{r}
library(tidyverse)
library(dplyr)
library(mice)

coffee <- read_csv("coffee.sort.csv") 

# coffee <- read_csv("coffee.sort.csv") %>%
#   mutate(across(where(is.character), as.factor)) #Needs updated dplyr

coffee[sapply(coffee, is.character)] <- lapply(coffee[sapply(coffee, is.character)], as.factor)  #  Converting all character to factors
```


### Missing data

```{r}
colSums(is.na(coffee))
``` 


#### Missing data in the country of origin

There is one NA value in the country of origin.
This coffee does have an owner and certification body that both only operate in Colombia according to their websites. Also all coffees in the data with those owners and that certification body are Colombian. Therefore, we assign this missing value to be Colombia.

```{r}
which(is.na(coffee$Country.of.Origin)) 

coffee[1198,]

coffee %>%
  filter(Owner == "racafe & cia s.c.a" & Certification.Body == "Almacafé") %>%
  count(Country.of.Origin)

coffee$Country.of.Origin[1198] <- "Colombia"

```

#### Delete erroneous data points

In the ranking categories, all values lie between at least 5 and 10, except for one coffee, which has 0 in all values. To delete this erroneous value, we keep all coffees that have at least one taste metric above 5.

```{r}
coffee %>% 
  select(Aroma:Uniformity) %>% 
  summary()

coffee %>%
  filter(Aroma < 5 | Flavor < 5 | Aftertaste < 5 | Acidity < 5 | Body < 5 | Balance < 5 | Uniformity < 5)



coffee <- 
  coffee %>% 
  filter(Aroma > 5 | Flavor > 5 | Aftertaste > 5 | Acidity > 5 | Body > 5 | Balance > 5 | Uniformity > 5)



coffee %>% 
  select(Aroma:Uniformity) %>% 
  summary()


  #  Prior to imputation we remove the outliers as our imputed values might be affected. With predictive mean matching 
  #   the imputed values always fall within the observed range, and thus these outliers might affect the imputations. Altitudes of > 10000m seem highly  unlikely
summary(coffee$altitude_mean_meters)
coffee <- filter(coffee, 
                 altitude_mean_meters < 10000.000 | is.na(altitude_mean_meters)) #Keep proper values and NAs

```


#### Imputation of larger amounts of missing data

The selected columns with missing data includes quakers, altitude_mean_meters, color, and processing method.
The number of complete cases in our selected coffee dataset is 930 (down from 1334).
Complete case analysis decreases our sample size, and can potentially lead to reduced statistical efficiency by increasing the potential for bias. Another option is multiple imputation, as our sample size will be preserved
We choose to handle our missing data problem by using multiple imputation.

```{r}
coffee_select <- coffee[c(1, 2, 12, 19:35)]  #  These variables seem most interesting and applicable for imputation
colnames(coffee_select)

fluxplot(coffee_select, labels = TRUE, main = "Outflux vs Influx", eqscplot = FALSE) # This plot indicates which variables to use

set.seed(1000)
ini <- mice(coffee_select, maxit=0, print=F)
pred <- ini$pred

pred[ "altitude_mean_meters", c("Processing.Method", "Color")] <- 0
pred[ "Processing.Method", c("altitude_mean_meters", "Color") ] <- 0
pred[ "Quakers", c("altitude_mean_meters", "Processing.Method", "Color")  ] <- 0
pred[ "Color" , c("altitude_mean_meters", "Processing.Method") ] <- 0

meth <- ini$method  #  Predictive mean matching for continous variables and a classification technique for the categorical variables
meth                  

imp <- mice(coffee_select, pred = pred, meth = meth, seed = 999, print=F, maxit = 10)  #  The multiple imputed datasets
stripplot(imp, altitude_mean_meters~.imp, pch=20, cex=1.4)
stripplot(imp, Processing.Method~.imp, pch=20, cex=1.4)
stripplot(imp, Quakers~.imp, pch=20, cex=1.4)
stripplot(imp, Color~.imp, pch=20, cex=1.4) #  Imputations seem to be adequate


coffee_select <- complete(imp)
sum(is.na(coffee_select))
```



### Construction of a 'continent' variable

```{r}
North_America <- c("United States", "United States (Hawaii)", "Mexico", "United States (Puerto Rico)", "Haiti")
Central_America <- c("Guatemala", "Costa Rica", "Honduras",  "Nicaragua","Panama", "El Salvador")
South_America <- c("Brazil", "Peru", "Colombia", "Ecuador")
Africa <- c("Ethiopia", "Uganda", "Tanzania, United Republic Of", "Kenya", "Burundi", "Rwanda", "Malawi", "Zambia", "Mauritius", "Cote d?Ivoire")
Asia <- c("Indonesia", "China", "Taiwan", "Thailand", "Papua New Guinea", "Vietnam", "Philippines", "Laos", "Myanmar", "India", "Japan")

attach(coffee_select)
coffee_select$Continent[Country.of.Origin %in% Africa]          <- "Africa"
coffee_select$Continent[Country.of.Origin %in% Asia]            <- "Asia"
coffee_select$Continent[Country.of.Origin %in% Central_America] <- "Central America"
coffee_select$Continent[Country.of.Origin %in% North_America]   <- "North America"
coffee_select$Continent[Country.of.Origin %in% South_America]   <- "South America"
detach(coffee_select)
coffee_select$Continent <- factor(coffee_select$Continent)

colnames(coffee_select[5:10])
coffee_select$mean_quality <- round((coffee_select$Aroma + coffee_select$Flavor + coffee_select$Aftertaste + coffee_select$Acidity + coffee_select$Body + coffee_select$Balance) / 6, 2)

save(coffee_select, file = "coffee_select.Rdata")

```


